```{r include=FALSE}
require(rethinking)
require(tidyverse)
```

## Exercice 1
Suppose the globe tossing data (Lecture 2, Chapter 2) had turned out to be 4
water and 11 land. Construct the posterior distribution.

```{r}
size <- 100
n_water <- 4
n_land <- 11
n_tosses <- n_water + n_land

p_grid <- seq(from = 0, to = 1, length.out = size)

prior <- rep(1, size)

likelihood <- dbinom(n_water, size = n_tosses, prob = p_grid)

unstd_post <- likelihood * prior

posterior <- unstd_post / sum(unstd_post)

tibble(p = p_grid, post = posterior) %>%
  ggplot(aes(x = p_grid, y = posterior)) +
  geom_line() + geom_point() +
  xlab("probability of water") +
  ylab("posterior probability") +
  labs(subtitle = str_c(as.character(size), " examples"))
```


 ## Exercice 2:
 Using the posterior distribution from 1, compute the posterior
 predictive distribution for the next 5 tosses of the same globe. I recommend
 you use the sampling method.

```{r}
n_simulations <- 1e4
sample_size <- 5

post_samples <- sample(
   p_grid,
   size = n_simulations,
   replace = TRUE,
   prob = posterior
)

w_pred <- rbinom(
  n_simulations,
  size = sample_size,
  prob = post_samples
)

tibble(w_pred) %>%
  ggplot(aes(x = w_pred)) +
  geom_histogram(bins = 18) +
  xlab("number of predicted water samples") +
  ylab("Frequency") +
  labs(title = "Predictive posterior distribution")
```

## Exercice 3
Use the posterior predictive distribution from 2 to calculate the probability of 3 or more water samples in the next 5 tosses.

```{r}
sum(w_pred >= 3) / n_simulations
```


## Exercice 4
This problem is an optional challenge for people who are
 taking the course for a second or third time. Suppose you observe W = 5
 water points, but you forgot to write down how many times the globe was
 tossed, so you don’t know the number of land points L. Assume that p = 0.7
 and compute the posterior distribution of the number of tosses N. Hint: Use
 the binomial distribution.
 
### Answer

$Pr(p_N | W, p_W, N) = \frac{Pr(W,p_W,N|p_N)*Pr(p_N)}{P(W,p_W,N)}$

#### Generative model

The function `sim_Tn_rounds_of_tosses` will simulate `Tn` (default to $1.000$)
rounds of globe tosses. Each round might consist of 1 to `n` (default to $30$).
Each toss has the probability `p` (default to $0.7$) of returning water.
The output is a `tibble` containing each round with its number of tosses `N` and
count of waters `W`.

```{r}
sim_Tn_rounds_of_tosses <- function(p=0.7, Tn=1e3, n=30) {
    N <- sample(1:n, size=Tn, replace=TRUE)
    W <- sapply(N, function(x) sum(rbinom(x, 1, p)))
    tibble(N,W)
}
```

Extracting data simulated with the default values (for sanity sake, we set the
seed to `120`):

```{r}
set.seed(120)
sim_data <- sim_Tn_rounds_of_tosses()
sim_data
sim_data %>%
    ggplot(aes(x=N, y=W)) +
    geom_jitter(alpha=0.4) +
    xlab("N of tosses") +
    ylab("Count of water (W)") +
    labs(title="Simulated 1.000 rounds of globe tosses")
```

a